\documentclass{article}

% Language setting
% Replace `english' with e.g. `spanish' to change the document language
\usepackage[english]{babel}

% Set page size and margins
% Replace `letterpaper' with `a4paper' for UK/EU standard size
\usepackage[letterpaper,top=2cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

% Useful packages
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}

\title{Campbell Thomasson Introduction for PS1}
\author{Campbell Thomasson}

\begin{document}
\maketitle

\section{Introduction}

Hello! My name is Campbell Thomasson. I am a first-year accounting Ph. D. student here at the University of Oklahoma. My primary research interests are tax policy, tax financial reporting, and tax avoidance. As far as economics goes, that was my other degree (besides accounting) when I went to my undergrad at Baylor University. My economics background has served me well in learning to conduct accounting research thus far (particularly regarding econometrics and capital markets). I want to take this class, because I believe it will make me a more efficient coder and manager of large sets of data, which would alsl be my main two goals for this class. For my project, I preliminary want to look at whether C-Corporations artificially managed their taxable income downward in 2020 due to the CARES Act, which allowed the companies to carry back any 2020 NOL up to five years to offset prior taxable income and receive a refund on that offset prior year taxable income, though I reserve the right to change my mind and choose another topic depending on feasibility of collecting the necessary data. After graduation, I plan to obtain a research position at an academic institution to both continue research and teach university students.

I am looking forward to class this semester!

\section{Equation}

\(a^2 + b^2 = c^2\)

\end{document}